---
title: "STATS 506 Final Project"
author: "Chuwen Li"
date: "12/7/2020"
output: 
  html_document: 
    theme: journal
    toc: yes
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

source('./analysis.R')
```

# <span id="jump">Introduction</span>
There are lots of myth around breastfeeding. How it can prevent children 
from asthma and overweight, improve children intelligence, or reduce the 
risk of children infections and hospitalizations. Even though nowadays 
there is a growing trend in formula feeding as the nutritional composition of
infant formula is increasingly close to that of breast milk and more young 
mothers get depressed from breastfeeding pain, lots of parents are still 
worries or blamed by their parents for risk infants' health by choosing formula
feeding. This project is aim to answer the question: 
 **Does breastfeeding prevent early childhood overweight?**

The analysis is based on fitting linear models with body measure index (BMI) 
as the dependent variable, and the explanatory factors as covariates, 
including sex, ethnicity, ever breastfed, mother smoked during pregnancy. The 
target group is 3 to 6 years old children.

# Data/Methods

The data used is from the National Health and Nutrition Examination Survey 
([NHANES](https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx?BeginYear=2017)). 
Specific datasets used are: Demographic data on the participants, 
health and nutrition examination survey on the participants, 
the answers to a diet behavior questionnaire, the answers to a early 
childhood questionnaire.

All participants in 2017 to 2018 continuous NHANES cycles whose age was 
between 3 to 6 years, who had complete demographic information on sex and 
ethnicity are first considered. BMI was measured all participants 2 year of 
age and older who attended a physical exam. Breast feeding history is the only 
variable takes participants up to 6 years of age and the question "Was {SP} 
ever breastfed or fed breastmilk?" was being asked. Mothers' smoking history 
during pregnancy is taken from the early childhood questionnaire. All 
the data values are adjusted based on two-year MEC exam weight.

Analysis was performed using "survey" package in R, which involves primary 
sampling unit, strata, and survey weights adjustment. The code for the analysis 
is available [here](https://github.com/lixx4228/Stats506_public). Two types of
linear models were considered. First one is taking the predict variable as 
continuous outcome. The second model uses a newly created binary variable: 
overweight as the predictor, which takes BMI greater and equal to 90th 
percentile as 1, and 0 otherwise. Both models were fitted with age, sex,
ethnicity, mother ever smoked during pregnancy, and ever being breastfed. Below
shows simple means for each explanatory variable with BMI.

```{r figs, echo=FALSE, fig.cap=cap1a, warning=FALSE, fig.height=5.6, fig.width=9}
cap1a = paste0("**Figure 1.** *BMI against age, sex, ethnicity, and ", 
               "momther smoked during pregnancy between breastfed and", 
               "non-breastfed children.*")

gridExtra::grid.arrange(p1, p2, p3, p4, ncol=2)
```


# Results

Table 1 displays the summary results for OLS and logistic regression models.
All models show a negative association between breastfed and being overweight 
or BMI gain, of which parameters are statistically significant for 95%
confidence level. There is a decrease risk of being overweight for ever 
breastfed children and based on the odds ratio, the risk is around 0.5 times 
smaller (95% CI, 0.37 - 0.75). The OLS model gives the result of 0.7 
less BMI value for a ever breastfed child compared to a never breastfed child,
holding all other variables constant. Age is positively associate with risk of
being overweight but only shows significance in the logistic model. Ethnicity
of Asia is the strongest predictor of child overweight and Hispanic is the only
ethnicity here that is positively associated with risk of overweight. Sex is 
not a good predictor nor as birth weight, which only shows the significance of 
first knots of cubic function in the OLS model. 

Within the sub sets of whether mother smoked during pregnancy, results were 
found that breastfed children whose mother did not smoke during pregnancy is 
more likely to have lower risk of overweight and the risk reduction is almost
doubled.  



# Conclusion

# To-do:
- prune codes
- add writing


```{r tab1, fig.cap = cap_tab1, echo=FALSE, warning=FALSE, message=FALSE}
cap_tab1 = "**Table 1.** *Statistical summary of all models.*"

tab0
```

-----

```{r tab2, fig.cap = cap_tab1, echo=FALSE, warning=FALSE}
cap_tab1 = "**Table 1.** *Statistical summary of between subsets.*"

tab4
```
