---
title: "STATS 506 Final Project"
author: "Chuwen Li"
date: "12/15/2020"
output:
  html_document:
    theme: journal
    toc: no
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sjPlot)

source('./analysis.R')
```

# <span id="jump">Introduction</span>
Nowadays there is a growing trend in formula feeding as the nutritional 
composition of infant formula is increasingly close to that of breast milk 
and more young mothers get depressed from breastfeeding pain. Still, lots of 
parents are worried or blamed by their parents for risking infants' health 
by only choose formula feeding. There is no doubt that many merits associated
with breastfeeding, including reducing risk of asthma and infections. Some 
of which are scientifically proved while others are widely various in opinions.
This project aims to answer one of the widespread breastfeeding questions: 

 **Does breastfeeding prevent early childhood overweight?**

The analysis is based on fitting linear models with body measure index (BMI) 
as the dependent variable, and the explanatory factors as covariates, 
including sex, ethnicity, ever breastfed, mother smoked during pregnancy. The 
target group is 3 to 6 years old children.

# Data/Methods

The data used is from the National Health and Nutrition Examination Survey 
([NHANES](https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx?BeginYear=2017)). 
Specific data sets used are: Demographic data on the participants, 
health and nutrition examination survey on the participants, 
the answers to a diet behavior questionnaire, the answers to a early 
childhood questionnaire.

The analysis was performed using the "survey" package in R, which involves 
primary sampling unit, strata, and survey weights adjustment. The code for 
the analysis is available [here](https://github.com/lixx4228/Stats506_public). 
Two types of linear models were considered. The first one was taking the 
predict variable as a continuous outcome. The second model used a newly 
created binary variable: 'overweight' as the predictor, which took BMI greater 
and equal to 90th percentile as 1, and 0 otherwise. Both models were fitted 
with age, sex, ethnicity, mother ever smoked during pregnancy, and ever being 
breastfed. Figure 1 shows simple means for each explanatory variable with BMI.

```{r figs, echo=FALSE, fig.cap=cap1a, fig.height=5.6, fig.width=9.5}
cap1a = paste0("**Figure 1.** *BMI against age, sex, ethnicity, and ", 
               "momther smoked during pregnancy between 3-to-6-year-old ",
               "breastfed and non-breastfed children, 2017-2018.*")

gridExtra::grid.arrange(p1, p2, p3, p4, ncol=2)
```


# Results

Table 1 displays the summary results for logistic regression model and OLS
model. For the OLS model, a natural spline with 2 knots was applied to 
the 'Birth weight' variable. Both models showed a negative association 
between breastfed children and being overweight or BMI gains with those younger 
than 6, of which parameters are statistically significant for 95% confidence 
level. There was a reduced risk of being overweight for breastfed 
children and based on the odds ratio, the risk was around 0.5 times smaller 
(95% CI, 0.37 - 0.75). The OLS model gave a result of 0.7 less BMI value 
for an ever breastfed child compared to a never breastfed child, holding all 
other variables constant. For other variables, age was positively associated 
with overweight in the logistic model but it was not significant in the OLS
model. In the data, no Asian ethnicity children whose BMI exceeded
90 percentile.

Breastfed children whose mothers never smoked during pregnancy were 
associated with a lower risk of overweight (OR 0.387) and the risk difference 
is almost doubled. Children in sex, age, and ethnicity subset behaves 
the same as the whole group.


# Conclusion

National data obtained from children (3-6 years) in 2017 and 2018 show 
ever breastfed children are less likely to experience early childhood 
overweight (BMI exceeds 90 percentile) compared to children who are only fed 
with formula. This result, however, does not entirely consistent with previous 
overweight and breastfeeding association study on 1998-1994 NHANES data 
(Hediger et al., 2001). Hediger, et. al. used BMI exceeded 95 percentile as 
the indicator of overweight and found no reduced risk of being overweight
between breastfed and non-breastfed children. Due to the technical difficulty 
of combining survey weights and mixed models in R, my study is limited to 
simple linear models. Thus, future application of mixed model to this data set
is a field of interest.




```{r tab1, echo=FALSE, warning=FALSE, message=FALSE}
cap_tab1 = paste0("**Table 1.** *Statistical summary of logistic model ",
                  "and OLS model.*")

tab_model(bmod3, smod1, digits = 3,
          pred.labels = c('(Intercept)', 'Breastfed[Yes]', 
                          'Birth weight, pounds', 'Age, years', 
                          'Sex[Male]', 'Ethnicity[Hispanic]', 
                          'Ethnicity[White]', 'Ethnicity[AfricanAmerican]',
                          'Ethnicity[Asian]', 'Ethnicity[Other]', 
                          'Mother smoked when pregnant[Yes]', 
                          'ns(Birth weight, 2)[1]', 'ns(Birth weight, 2)[2]'), 
          dv.labels = c('Overweight<br>(BMI \u2265 90th Percentile)',
                        'BMI<br>(kg/m²)'), 
          title = cap_tab1)
```

---

# Reference

[1] Hediger ML, Overpeck MD, Kuczmarski RJ, Ruan WJ. 
Association Between Infant Breastfeeding and Overweight in Young Children. 
*JAMA*. 2001;285(19):2453–2460. doi:10.1001/jama.285.19.2453

